Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants\thanks{Supported by EPSRC grant EP/J012564/1, ERC project 280053 (CPROVER) and the H2020 FET OPEN 712689 SC$^2$.}

### 2017-05-04 cav2017/cav2017.tex
sec 2

%-------------------------------
\section{Related work}
\label{sec:relw}
%-------------------------------

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop system
\eqref{eq:closedloopss} meets given safety specifications.  A safety
specification gives raise to a requirement on the states of the model, such
that the feedback controller (namely the choice of the gains matrix~$K$)
must ensure that the state never violates the requirement.  Note that a
stable, closed-loop system is not necessarily a safe system: indeed, the
state values may leave the safe part of the state space while they converge
to the equilibrium, which is typical in the case of oscillatory dynamics. 
However, a safe system will always be stable.  In~this work, the safety
property is modeled as follows:
%
\begin{equation}
\label{eq:safetyliteral}
\phi_\mathit{safety}\iff \forall k\ge 0.\, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
%
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds
for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th
instant, respectively.  This means that the states will always be within the
$n$-dimensional hyper-box.

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds:
$\phi_\mathit{input} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} $, 
$\phi_\mathit{init} = \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$
For the former, this means that the control input saturates in view of
physical constraints.

@@

The models we consider have two sources of error due to numeric
representation.  The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$ and $x_k$.  The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
fixed-point representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{appendix:numerical_errors}.

@@

We will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real
number $x$ represented at the fixed-point precision of the controller, and
$\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$
represented at the fixed-point precision of the plant model ($I_p \geq I_c
\wedge\allowbreak F_p \geq F_c$).  Thus any mathematical operations in the
digital controller will be in the range of $\mathcal{F}_{\langle I_c,F_c
\rangle}$, and all other calculations will be carried out in the range of
$\mathcal{F}_{\langle I_p,F_p \rangle}$.  The physical plant operates in the
reals, which means our verification phase must account for calculation
errors in the model.

@@

This induces two types of the errors detailed above: first, the truncation
error due to representing $x_k$ as $\mathcal{F}{\langle I_c,F_c
\rangle}(x_{k})$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
noise.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and
zeros positions, degrading the closed-loop dynamics, causing steady-state
errors (see Appendix~\ref{sec:appendix:LTIbackground} for details) and
eventually destabilizing the system~\cite{Bessa16}.  However, since in this
paper we require stability only as a precursor to safety, it is sufficient
to check that the (perturbed, noisy) model converges to a neighborhood of
the equilibrium within the safe set.

@@

In this section, we describe our technique for synthesizing safe digital
feedback controllers using CEGIS.  For this purpose, we first provide the
synthesizer's general architecture, followed by describing our two
approaches to synthesizing safe controllers: the first one is a baseline
approach that relies on a na\"ive unfolding of the transition relation,
whereas the second uses abstraction to evaluate all possible progressions of
the system.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Naive Approach: CEGIS with Multi-staged Verification}
\label{sec:CEGIS-precision-incrementation}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

An overview of the algorithm for controller synthesis is given in
Fig.~\ref{fig:CEGIS-precision-increment}.  One important observation is that
we verify and synthesize a controller over $k$ time steps.  We~then compute
a completeness threshold~\cite{DBLP:conf/vmcai/KroeningS03}, $\overline{k}$,
for this controller, and verify correctness for $\overline{k}$ time steps. 
Essentially, $\overline{k}$ is the number of iterations required to
sufficiently unwind the closed-loop state-space system such that the
boundaries are not violated for any $k{>}\overline{k}$.

\begin{lemma} There exists a $\overline{k}$ such that it is
sufficient to unwind the closed-loop state-space system up to $\overline{k}$
in order to ensure that $\phi_\mathit{safety}$ holds. 
\end{lemma}

\begin{proof}
%
A \blue{stable} control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this is not the case in general when dealing with
complex eigenvalues.  Consider the closed-loop matrix that updates the
states in every discrete time step, and select the eigenvalue $\vartheta$
with the smallest (non-trivial) imaginary value.  Between every pair of
consecutive time steps $k\,T_s$ and $(k+1)\,T_s$, the dynamics projected on
the corresponding eigenspace rotate $\vartheta T_s$ radians.  Thus, taking
$\overline{k}$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after
$k{\geq}\overline{k}$ steps we have completed a full rotation, which results
in a point closer to the origin.  This is the completeness threshold.

%
\end{proof}

@@

\item The inductive synthesis phase ({\sc synthesize}) uses BMC to
compute a candidate solution $K$ that satisfies both the stability criteria
(Sec.~\ref{ssec:stability}) and the safety specification
(Sec.~\ref{ssec:safety}).  To synthesize a controller that satisfies the
stability criteria, we require that a computed characteristic polynomial
satisfies Jury's criteria.  The details of this calculation can be found in
the Appendix.

@@

\item The first {\sc verify} stage, {\sc safety}, checks that candidate
solution $K$, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps where we assume $k$ to be under the
completeness threshold.  After unfolding the transition system corresponding
to the previously synthesized controller $k$ steps, we check that the safety
specification holds for any initial state. This is shown in Alg.~~\ref{alg:safetycheck}.

@@

\begin{lemma} 
If a controller is safe for each of the corner cases of our hypercube of
allowed initial states, it is safe for any initial state in the
hypercube. 
\end{lemma}

Thus we only need to check $2^n$ initial states, where $n$
is the dimension of the state space (number of continuous variables). 
\begin{proof}
Consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i 
      = \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
%
where $x_k^i$ denotes the trajectories obtained from the single vertex
$v_i$.  We~conclude that any $k$-step trajectory is encompassed, within a
convex set, by those generated from the vertices.
\end{proof}

@@

%-------------------------------
\subsection{Description of the Benchmarks}
\label{experimental-setup}
%-------------------------------

@@

\begin{table}
\centering
%\scriptsize
\begin{tabular}{| r | l | c | c | r | c | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{Order} & \multicolumn{2}{|c|}{Multi-staged}                 & \multicolumn{2}{|c|}{Abstraction} \\
   &                                  & & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} \\\hline
1  & Cruise Control  & 1 & 8,16   & 8.40\,s & 16,16  &   2.17\,s \\
2  & DC Motor          & 2 & 8,16   & 9.45\,s & 20,20  &   2.06\,s \\
3  & Helicopter        & 3 & \xmark & \xmark  & 16,16  &   1.37\,s \\
4  & Inverted Pendulum & 4 & 8,16   & 9.65\,s & 16,16  &   0.56\,s \\
5  & Magnetic Pointer  & 2 & \xmark & \xmark  & 28,28  &  44.14\,s \\
6  & Magnetic Suspension & 2 & 12,20  &10.41\,s & 16,16  &   0.61\,s \\
7  & Pendulum          & 2 & 8,16   &14.02\,s & 16,16  &   0.60\,s \\
8  & Suspension        & 2 & 12,20  &73.66\,s & \xmark &    \xmark \\
9  & Tape Driver       & 3 & 8,16   &10.10\,s & 16,16  &  68.24\,s \\
10 & Satellite          & 2 & 8,16   & 9.43\,s & 16,16  &   0.67\,s \\\hline
%
\end{tabular}
\caption{Experimental results\label{tab:results}}
\end{table}

@@

%-------------------------------
\section{Stability of closed-loop models}
\label{sec:appendix-stability}
%-------------------------------

@@

%-------------------------------
\subsection{Stability of Closed-loop Systems with Fixed-point Controller Error}
%-------------------------------

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{LTI system Transformations and Equivalences} \label{sec:appendix:LTIbackground}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

>>>>

%-------------------------------
\section{Related Work}
\label{sec:relw}
%-------------------------------

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop system
\eqref{eq:closedloopss} meets given safety specifications.  A safety
specification gives raise to a requirement on the states of the model, such
that the feedback controller (namely the choice of the gains matrix~$K$)
must ensure that the state never violates the requirement.  Note that a
stable, closed-loop system is not necessarily a safe system: indeed, the
state values may leave the safe part of the state space while they converge
to the equilibrium, which is typical in the case of oscillatory dynamics. 
%However, a safe system will always be stable.  
In~this work, the safety property is expressed as follows:
%
\begin{equation}
\label{eq:safetyliteral}
\phi_\mathit{safety}\iff \forall k\ge 0.\, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
%
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds
for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th
instant, respectively.  This means that the states will always be within an $n$-dimensional hyper-box.

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds:
$\phi_\mathit{input} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} $, 
$\phi_\mathit{init} = \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$
For the former, this means that the control input might saturates in view of
physical constraints.

@@

The models we consider have two sources of error that are due to numerical  
representation.  The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$ and $x_k$.  The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
fixed-point representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{appendix:numerical_errors}.

@@

We will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real
number $x$ represented at the fixed-point precision of the controller, and
$\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$
represented at the fixed-point precision of the plant model (\blue{[justify this assumption:]} $I_p \geq I_c$ and $\allowbreak F_p \geq F_c$).  Thus any mathematical operations in the
digital controller will be in the range of $\mathcal{F}_{\langle I_c,F_c
\rangle}$, and all other calculations will be carried out in the range of
$\mathcal{F}_{\langle I_p,F_p \rangle}$.  
\blue{[this statement is potentially confusing: are we referring to quantisation effects, or to the uncertainty errors that we have handled in our HSCC17 work?]
The physical plant operates in the
reals, which means our verification phase must account for calculation
errors in the model. 
}

@@

This induces two types of the errors detailed above: first, the truncation
error due to representing $x_k$ as $\mathcal{F}{\langle I_c,F_c
\rangle}(x_{k})$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
additive noise, which might de-stabilize the closed-loop model.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and
zeros positions, degrading the closed-loop dynamics, causing steady-state
errors (see Appendix~\ref{sec:appendix:LTIbackground} for details) and
eventually de-stabilizing the system~\cite{Bessa16}.  However, since in this
paper we require stability only as a precursor to safety, it is sufficient
to check that the (perturbed, noisy) model converges to a neighborhood of
the equilibrium within the safe set.

@@

In this section, we describe our technique for synthesizing safe digital
feedback controllers using CEGIS.  For this purpose, we first provide the
synthesizer's general architecture, followed by describing our two
approaches to synthesizing safe controllers: the first one is a baseline
approach that relies on a na\"ive unfolding of the transition relation,
whereas the second uses abstraction to evaluate all possible executions of
the system.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Na\"ive Approach: CEGIS with multi-staged verification}
\label{sec:CEGIS-precision-incrementation}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

An overview of the algorithm for controller synthesis is given in
Fig.~\ref{fig:CEGIS-precision-increment}.  One important observation is that
we verify and synthesize a controller over $k$ time steps.  We~then compute
a completeness threshold $\overline{k}$~\cite{DBLP:conf/vmcai/KroeningS03}
for this controller, and verify correctness for $\overline{k}$ time steps. 
Essentially, $\overline{k}$ is the number of iterations required to
sufficiently unwind the closed-loop state-space system, 
which ensures that the boundaries are not violated for any other $k{>}\overline{k}$.

\begin{theorem} There exists a finite $\overline{k}$ such that it is
sufficient to unwind the closed-loop state-space system up to $\overline{k}$
in order to ensure that $\phi_\mathit{safety}$ holds. 
\end{theorem}

\begin{proof}
%
A stable control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this is not the case in general when dealing with
complex eigenvalues.  Consider the closed-loop matrix that updates the
states in every discrete time step, and select the eigenvalue $\vartheta$
with the smallest (non-trivial) imaginary value.  Between every pair of
consecutive time steps $k\,T_s$ and $(k+1)\,T_s$, the dynamics projected on
the corresponding eigenspace rotate $\vartheta T_s$ radians.  Thus, taking
$\overline{k}$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after
$k{\geq}\overline{k}$ steps we have completed a full rotation, which results
in a point closer to the origin.  The synthesized $\overline{k}$ is the completeness threshold.
\qed 
%
\end{proof}

@@

\item The inductive synthesis phase ({\sc synthesize}) uses BMC to
compute a candidate solution $K$ that satisfies both the stability criteria
(Sec.~\ref{ssec:stability}) and the safety specification
(Sec.~\ref{ssec:safety}).  To synthesize a controller that satisfies the
stability criteria, we require that a computed polynomial
satisfies Jury's criterion~\cite{fadali}.  The details of this calculation can be found in
the Appendix.

@@

\item The first {\sc verify} stage, {\sc safety}, checks that the candidate
solution $K$, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps where we assume $k$ to be under the
completeness threshold.  After unfolding the transition system corresponding
to the previously synthesized controller $k$ steps, we check that the safety
specification holds for any initial state. This is shown in Alg.~~\ref{alg:safetycheck}.

@@

\begin{theorem} 
If a controller is safe for each of the corner cases of our hypercube of
allowed initial states, it is safe for any initial state in the hypercube. 
\end{theorem}

Thus we only need to check $2^n$ initial states, where $n$
is the dimension of the state space (number of continuous variables). 
\begin{proof}
Consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i 
      = \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
%
where $x_k^i$ denotes the trajectories obtained from the single vertex
$v_i$.  We~conclude that any $k$-step trajectory is encompassed, within a
convex set, by those generated from the vertices. 
\qed
\end{proof}

@@

%-------------------------------
\subsection{Description of the benchmarks}
\label{experimental-setup}
%-------------------------------

@@

\begin{table}
\centering
%\scriptsize
\begin{tabular}{| r | l | c | c | r | c | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{Order} & \multicolumn{2}{|c|}{Multi-staged}                 & \multicolumn{2}{|c|}{Abstraction} \\
   &                                  & & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} \\\hline
1  & Cruise Control  & 1 & 8,16   & 8.40\,s & 16,16  &   2.17\,s \\
2  & DC Motor          & 2 & 8,16   & 9.45\,s & 20,20  &   2.06\,s \\
3  & Helicopter        & 3 & \xmark & \xmark  & 16,16  &   1.37\,s \\
4  & Inverted Pendulum & 4 & 8,16   & 9.65\,s & 16,16  &   0.56\,s \\
5  & Magnetic Pointer  & 2 & \xmark & \xmark  & 28,28  &  44.14\,s \\
6  & Magnetic Suspension & 2 & 12,20  &10.41\,s & 16,16  &   0.61\,s \\
7  & Pendulum          & 2 & 8,16   &14.02\,s & 16,16  &   0.60\,s \\
8  & Suspension        & 2 & 12,20  &73.66\,s & \xmark &    \xmark \\
9  & Tape Driver       & 3 & 8,16   &10.10\,s & 16,16  &  68.24\,s \\
10 & Satellite          & 2 & 8,16   & 9.43\,s & 16,16  &   0.67\,s \\\hline
%
\end{tabular}
\vspace{0.05in}
\caption{Experimental results\label{tab:results}}
\end{table}

@@

%-------------------------------
\section{Stability of Closed-loop Models}
\label{sec:appendix-stability}
%-------------------------------

@@

%-------------------------------
\subsection{Stability of closed-loop models with fixed-point controller error}
%-------------------------------

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Errors in LTI models} \label{sec:appendix:LTIbackground}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
